---
title: "VB R_Assignment"
author: "Vishesh Bhatia"
date: "3/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
setwd('.')
```

## Reading the Genotypes and SNPs file
Through the code below, we are able to load the files fang_et_al_genotypes.txt 
and snp_position file into our global environment tab. 
```{r}
genotypes <- read.table("https://raw.githubusercontent.com/EEOB-BioData/BCB546-Spring2021/main/assignments/UNIX_Assignment/fang_et_al_genotypes.txt", header = TRUE, sep = "\t")
SNPs <- read.table("https://raw.githubusercontent.com/EEOB-BioData/BCB546-Spring2021/main/assignments/UNIX_Assignment/snp_position.txt", header = TRUE, sep = "\t")
```

# Part 1
## DATA INSPECTION

We were able to analyze both the genotypes and SNP files and the code for that is below. 

```{r}
dim(genotypes)
#Provided us with the dimensions of the file: 2782 rows and 986 columns
dim(SNPs)
#Provided us with the dimensions of the file: 983 rows and 15 columns

genotypes[1:10, 1:10]
SNPs[1:10, 1:10]
#Prints the first 10 rows and columns and helps us get an idea of what the file looks like. 

class(genotypes)
class(SNPs)
#Provides us the information about the vectors and looks like both files are data.frame format. 

str(genotypes)
str(SNPs)
#Provides information about structure of the data frame such as name, type and preview of data in each column. 

```
## PART 2
## Data Processing

Sort out the genotypes based on the Group and create separate data files of each species. 

```{r}
#Filter out the maize data
maize_g <- filter(genotypes, Group == 'ZMMIL' | Group == 'ZMMLR' | Group == 'ZMMMR')

#Filter out the Teosinte Data
teosinte_g <- filter(genotypes, Group == 'ZMPBA' | Group == 'ZMPIL' | Group == 'ZMPJA')
```

### Transposing the Maize Data

```{r}
maize_g <- column_to_rownames(maize_g, var = "Sample_ID")
transposed_maize <- t(maize_g) %>% as.data.frame() %>% rownames_to_column(., var = "SNP_ID") 
transposed_maize <- transposed_maize[3:nrow(transposed_maize),]
Mdata <- merge(SNPs, transposed_maize, by = "SNP_ID")
Mdata <- select(Mdata, SNP_ID, Chromosome, Position, everything())
```

### Tranposing the Teosinte Data

```{r}
teosinte_g <- column_to_rownames(teosinte_g, var = "Sample_ID")
transposed_teosinte <- t(teosinte_g) %>% as.data.frame() %>% rownames_to_column(., var = "SNP_ID") 
transposed_teosinte <- transposed_teosinte[3:nrow(transposed_teosinte),]
TData <- merge(SNPs, transposed_teosinte, by = "SNP_ID")
TData <- select(TData, SNP_ID, Chromosome, Position, everything())
```

## Maize Output

```{r}
chr_maize <- filter(Mdata, Chromosome != "multiple" & Chromosome != "unknown")
for (i in 1:length(unique(chr_maize$Chromosome))){
  chrm <-  chr_maize %>% filter(Chromosome == i) %>% arrange(Position)
  write.table(chrm, file = paste("./Maize/Maize_Part_A",i, sep="_"))
}

```

```{r}
Subbed_Maize <- as_tibble(lapply(Mdata, gsub, pattern ="?", replacement ="-", fixed = TRUE))
Subbed_Chr_Maize <- filter(Subbed_Maize, Chromosome != "multiple" & Chromosome != "unknown")
for (i in 1:length(unique(Subbed_Chr_Maize$Chromosome))) {
  chrm_subbed <- Subbed_Chr_Maize %>% filter(Chromosome == i) %>% arrange(desc(Position))
  write.table(chrm_subbed, file = paste("./Maize/Maize_Part_B", i, sep = "_"))
  }

```

## Teosinte Data
```{r}
chr_teosinte <- filter(TData, Chromosome != "multiple" & Chromosome != "unknown")
for (i in 1:length(unique(chr_teosinte$Chromosome))) {
  chrt <- chr_teosinte %>% filter(Chromosome == i) %>% arrange(Position)
  write.table (chrt, file = paste("./Teosinte/Teosinte_Part_A", i, sep = "_"))
}
```

```{r}
Subbed_Teosinte <- as_tibble(lapply(TData, gsub, pattern = "?", replacement = "-", fixed = TRUE))
Subbed_Chr_Teosinte <- filter(Subbed_Teosinte, Chromosome != "multiple" & Chromosome != "unknown")
for (i in 1:length(unique(Subbed_Chr_Teosinte$Chromosome))) {
    chrt_subbed <- Subbed_Chr_Teosinte %>% filter(Chromosome == i) %>% arrange(desc(Position))
    write.table(chrt_subbed, file = paste("./Teosinte/Teosinte_Part_B", i, sep = "_"))
}
```

# PART 3
# DATA VISUALIZATION


